---
title: "Filter tourists"
author: "Rujal Shrestha"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  html_document:
    theme: flatly
    highlight: zenburn
  # pdf_document:
  #   latex_engine: xelatex
---

```{r libraries, include=FALSE}
# Import Libraries
library(tidyverse)
library(readr)
```

### Load datasets

```{r, CACHE = FALSE}
secondary_purpose_of_visit <- read.csv(
  "./datasets/secondary_purpose_of_visit.csv",
  header = TRUE
)

View(secondary_purpose_of_visit)

survey_main_header <- read.csv(
  "./datasets/survey_main_header.csv",
  header = TRUE
)

View(survey_main_header)
```

`secondary_purpose_of_visit` has 22047 rows  
`survey_main_header` has 22360 rows

`survey_main_header` has dupliate rows

### Combine data frames

```{r, CACHE = FALSE}
combined_df <- inner_join(
  survey_main_header,
  secondary_purpose_of_visit,
  by = "response_id"
)

View(combined_df)
```

`combined_df` has 22360 rows, contains duplicate rows as well

### Identify and remove duplicate responses

```{r, CACHE = FALSE}
duplicate_rows <- combined_df |>
  count(response_id) |>
  filter(n > 1) |>
  select(response_id)

View(duplicate_rows)

filtered_combined_df <- combined_df |>
  filter(!response_id %in% duplicate_rows$response_id)

View(filtered_combined_df)
```
now `filtered_combined_df` has 21756 rows, which are all unique

### Extract purpose related columns

```{r, CACHE = FALSE}
purpose <- filtered_combined_df |>
  select(
    response_id,
    purpose_of_visit_main,
    purpose_subtype,
    reasons_for_travel_secondary
  )

View(purpose)
```

### Primary reasons for visit 

```{r, CACHE = FALSE}
purpose |>
  group_by(purpose_of_visit_main) |>
  summarise(n = n())
```

### Secondary reasons for visit

```{r, CACHE = FALSE}
purpose |>
  group_by(reasons_for_travel_secondary) |>
  summarise(n = n()) |>
  View()
```

### Filtered tourists response_ids

```{r, CACHE = FALSE}
tourist_response_ids <- purpose |>
  filter(
    reasons_for_travel_secondary == "Holiday / vacation" |
      purpose_of_visit_main == "Holiday / vacation"
  ) |>
  pull(response_id)

length(tourist_response_ids)
```

16,774 total unique tourist responses

### Combined table 

Can use `tourist_response_ids` to filter out non-tourists in any table.  
Example: `survey_main_header` below is filtered and results are stored in 
`tourist_survey_main_header`.

```{r, CACHE = FALSE}
tourist_survey_main_header <- survey_main_header |>
  filter(response_id %in% tourist_response_ids)

View(tourist_survey_main_header)
```

### Saving to csv file

```{r, CACHE = FALSE}
write_csv(
  tourist_survey_main_header,
  "./outputs/tourist_survey_main_header.csv"
)
```
