---
title: "K Means + Recommendation System"
author: "Yuxuan Wu"
date: "2025-04-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, include=FALSE}
# Import Libraries
library(tidyverse)
library(readr)
library(ggplot2)
library(scales)
library(dplyr)
library(fastDummies)

library(stringr)


```


### 1. Data Preparation
```{r}
# Filter Tourists with "Holiday / Vacation" as Primary or Secondary Travel Purpose
survey_main_header <- read.csv(
  "./dataset for modeling/survey_main_header.csv",
  header = TRUE
)

secondary_purpose_of_visit <- read.csv(
  "./dataset for modeling/secondary_purpose_of_visit.csv",
  header = TRUE
)

combined_df <- inner_join(
  survey_main_header,
  secondary_purpose_of_visit,
  by = "response_id"
)

duplicate_rows <- combined_df |>
  count(response_id) |>
  filter(n > 1) |>
  select(response_id)

filtered_combined_df <- combined_df |>
  filter(!response_id %in% duplicate_rows$response_id)

purpose <- filtered_combined_df |>
  select(
    response_id,
    purpose_of_visit_main,
    purpose_subtype,
    reasons_for_travel_secondary
  )

purpose |>
  group_by(purpose_of_visit_main) |>
  summarise(n = n())

purpose |>
  group_by(reasons_for_travel_secondary) |>
  summarise(n = n()) 

tourist_response_ids <- purpose |>
  filter(
    reasons_for_travel_secondary == "Holiday / vacation" |
      purpose_of_visit_main == "Holiday / vacation"
  ) |>
  pull(response_id)

tourist_survey_main_header <- survey_main_header |>
  filter(response_id %in% tourist_response_ids)
```



```{r}
# Filter tourist records in other tables based on selected response_ids, and merge into a consolidated table

# Select relevant columns from tourist_survey_main_header for modeling
tourist_survey_main_header <- tourist_survey_main_header |>
  select(response_id, age_range, country_of_residence, gender, no_days_in_nz)

# Load additional tables containing satisfaction and expenditure information
visitor_satisfaction <- read.csv(
  "./dataset for modeling/visitor_satisfaction.csv",
  header = TRUE
)

expenditure_by_industry <- read.csv(
  "./dataset for modeling/expenditure_by_industry.csv",
  header = TRUE
)

# Filter satisfaction and expenditure records corresponding to the selected tourist response_ids
tourist_visitor_satisfaction <- visitor_satisfaction |>
  filter(response_id %in% tourist_response_ids)

tourist_expenditure <- expenditure_by_industry |>
  filter(response_id %in% tourist_response_ids)

# Merge all datasets on response_id to create a unified tourist dataset
tourist_full <- tourist_survey_main_header |>
  left_join(tourist_visitor_satisfaction, by = "response_id") |>
  left_join(tourist_expenditure, by = "response_id")

```


```{r}
# 1. Select features relevant for modeling
tourist_full <- tourist_full |>
  select(response_id, age_range, country_of_residence, gender, no_days_in_nz,
         satisfaction_rating, cost_accomm, cost_dom_travel, cost_food_drink,
         cost_entertainment, cost_shopping, cost_other, cost_tour_package,
         cost_dom_flights, cost_car_rentals, cost_eating_out,
         cost_day_cruise, exc_cost_int_flights, exc_cost_multi_night_cruise,
         exc_cost_travel_insurance)

# 2. Encode categorical variables as factors
tourist_full <- tourist_full |>
  mutate(
    age_range = as.factor(age_range),
    gender = as.factor(gender),
    country_of_residence = as.factor(country_of_residence)
  )

# 3. Create the 'total_spending' feature by summing all relevant cost components
tourist_full <- tourist_full |>
  mutate(
    total_spending = rowSums(
      across(
        c(cost_accomm, cost_dom_travel, cost_food_drink,
          cost_entertainment, cost_shopping, cost_other,
          cost_tour_package, cost_dom_flights, cost_car_rentals,
          cost_eating_out, cost_day_cruise, exc_cost_int_flights,
          exc_cost_multi_night_cruise, exc_cost_travel_insurance),
        ~ replace_na(., 0)  # Replace missing cost values with zero before summing
      )
    )
  )

# 4. Remove records with missing values in essential fields
tourist_full <- tourist_full |>
  filter(
    !is.na(age_range),
    !is.na(country_of_residence),
    !is.na(gender),
    !is.na(no_days_in_nz),
    !is.na(total_spending),
    !is.na(satisfaction_rating)
  )

# 5. Transform 'satisfaction_rating' from text to numeric and derive a binary satisfaction label
tourist_full <- tourist_full |>
  mutate(
    satisfaction_rating = str_extract(satisfaction_rating, "^[0-9]") |> as.numeric(),
    satisfaction_binary = ifelse(satisfaction_rating %in% c(1, 2), "Low", "High"),
    satisfaction_binary = as.factor(satisfaction_binary)
  )

# 6. Finalize the dataset with selected features for modeling
model_data <- tourist_full |>
  select(
    response_id,
    age_range,
    country_of_residence,
    gender,
    no_days_in_nz,
    total_spending,
    satisfaction_binary
  )
```



## K-Means Recommendation System
```{r}
### 0. Load libraries

library(dplyr)
library(ggplot2)
library(caret)
library(cluster)   # For clustering visualization
library(factoextra) # For PCA visualization

# 1. Prepare the data

# Convert age_range to numeric midpoint
age_midpoints <- c(
  "20 - 24" = 22, 
  "25 - 29" = 27, 
  "30 - 34" = 32, 
  "35 - 39" = 37, 
  "40 - 44" = 42, 
  "45 - 49" = 47, 
  "50 - 54" = 52, 
  "55 - 59" = 57,
  "60 - 64" = 62, 
  "65 - 69" = 67, 
  "70 - 74" = 72, 
  "75 or older" = 77,  
  "Under 20" = 18,     
  "80 - 84" = 82,
  "85 - 89" = 87
)

# Remove age range of 'Rather not say'
model_data <- model_data |>
  filter(age_range != "Rather not say")

# Transform into numerical
model_data <- model_data |>
  mutate(
    age_numeric = as.numeric(age_midpoints[as.character(age_range)])
  )


# One-hot encode gender
cluster_features <- model_data |>
  select(age_numeric, no_days_in_nz, total_spending, gender) |>
  dummy_cols(select_columns = "gender", remove_first_dummy = TRUE, remove_selected_columns = TRUE)

# Standardize numeric features
cluster_features_scaled <- scale(cluster_features)


# 2. Perform K-Means clustering
set.seed(123)
kmeans_model <- kmeans(cluster_features_scaled, centers = 5, nstart = 25)

# Attach cluster labels to model_data
model_data$cluster <- kmeans_model$cluster

# 3. Link tourist activities
activities <- read.csv("./dataset for modeling/activities.csv")

tourist_activities <- activities |>
  filter(response_id %in% model_data$response_id) |>
  left_join(model_data[, c("response_id", "cluster")], by = "response_id")

# 4. Summarize top activities for each cluster
top_activities_per_cluster <- tourist_activities |>
  group_by(cluster, activity) |>
  summarise(count = n(), .groups = "drop") |>
  group_by(cluster) |>
  slice_max(order_by = count, n = 5)




```


```{r}
### 5. Visualize clustering results using PCA

# Perform Principal Component Analysis (PCA)
pca_result <- prcomp(cluster_features_scaled)

# Create a dataframe for plotting
pca_data <- as.data.frame(pca_result$x)
pca_data$cluster <- as.factor(model_data$cluster)

# Plot PCA results with cluster coloring
ggplot(pca_data, aes(x = PC1, y = PC2, color = cluster)) +
  geom_point(alpha = 0.7, size = 2) +
  labs(title = "Tourist Clusters Visualized via PCA",
       x = "Principal Component 1",
       y = "Principal Component 2") +
  theme_minimal() +
  scale_color_brewer(palette = "Set1")


```


Comment:


```{r}
### 6. Interpret clustering results

# Summarize average features per cluster
cluster_profiles <- model_data |>
  group_by(cluster) |>
  summarise(
    mean_age = round(mean(age_numeric), 1),
    mean_days = round(mean(no_days_in_nz), 1),
    mean_spending = round(mean(total_spending), 1),
    n = n()
  ) |>
  arrange(cluster)

print(cluster_profiles)


```


```{r}
### 7. Generate textual description for each cluster

generate_cluster_description <- function(cluster_row) {
  paste0(
    "Cluster ", cluster_row$cluster, ": On average, tourists are around ", 
    cluster_row$mean_age, " years old, stay for about ", 
    cluster_row$mean_days, " days, and spend approximately $", 
    cluster_row$mean_spending, ". (", cluster_row$n, " tourists)"
  )
}

# Output cluster profiles as text
apply(cluster_profiles, 1, function(row) generate_cluster_description(as.list(row)))

```


```{r}
### 8. Recommend Top-3 activities for each cluster (excluding 'None of these')

top_activities_per_cluster <- tourist_activities |>
  group_by(cluster, activity) |>
  summarise(count = n(), .groups = "drop") |>
  filter(activity != "None of these") |>   
  group_by(cluster) |>
  slice_max(order_by = count, n = 3)


```



```{r}
### 9. Predict cluster for a new tourist and recommend activities

# Define a new tourist's profile
new_user <- data.frame(
  age_numeric = 30,
  no_days_in_nz = 12,
  total_spending = 6000,
  gender_Male = 1,
  gender_Female = 0,
  gender_Another_Gender = 0  
)

# Standardize new user's features
new_user_scaled <- as.data.frame(scale(
  new_user,
  center = attr(cluster_features_scaled, "scaled:center"),
  scale = attr(cluster_features_scaled, "scaled:scale")
))


# Calculate Euclidean distances to each cluster center
distances <- apply(kmeans_model$centers, 1, function(center) sum((center - as.numeric(new_user_scaled))^2))

# Assign the new tourist to the closest cluster
predicted_cluster <- which.min(distances)

print(paste("New tourist predicted to belong to Cluster", predicted_cluster))

# Retrieve Top-3 recommended activities
recommended_activities <- top_activities_per_cluster |>
  filter(cluster == predicted_cluster)

# Output the recommendation explanation
cluster_profile_row <- cluster_profiles |> filter(cluster == predicted_cluster)

cat(
  "Recommendation Explanation:\n",
  "Based on the tourist's age, gender, spending, and days of stay, they are most similar to Cluster ",
  predicted_cluster, ".\n", 
  generate_cluster_description(as.list(cluster_profile_row)), "\n",
  "Recommended activities are:\n",
  paste0("- ", recommended_activities$activity, collapse = "\n")
)

```




```{r}
# Check the proportion of variance explained by each principal component of PCA
summary(pca_result)

```

```{r}
library(factoextra)

fviz_pca_var(pca_result,
             col.var = "contrib",  
             gradient.cols = c("lightblue", "blue", "darkblue"),
             repel = TRUE) +  
  labs(title = "Variable Contributions to Principal Components")

```

Based on PCA variable contribution analysis, the first principal component (PC1) predominantly captures gender-related variance, particularly distinguishing between male and female tourists. The second component (PC2) is largely driven by behavioral traits such as duration of stay and total spending, indicating a potential differentiation between short-term/low-spending and long-term/high-spending tourists.


```{r}

```

